Computational design of RNA-based oscillatory circuits

### 2015-08-23 MScWriteup.tex
writing the background section

\author{J.~Binysh
        \\ \IEEEmembership{* University of Warwick, Complexity Department}
        \vspace{-3em}}

@@

\begin{abstract}
the abstract
\end{abstract}

@@

\section{Introduction}
\label{sec: Intro}
The process of gene expression can be briefly summarised as follows: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}). This RNA molecule (known as messenger RNA, or mRNA) makes its way to a piece of cellular machinery called the Ribosome, which reads it, and makes a protein - which protein is made depends on the DNA sequence originally read (\textit{translation}).

 The path from genetic transcription to protein expression is naturally regulated in many ways \cite{MolecularBiology}. This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues. The natural cell machinery in place to perform this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.

@@

 \begin{itemize}
 \item goal of the project - emph on quantitative modelling of biological system and data, getting a quantitative model, finding its parameters.
 \item Biological background, taken from those 3 papers.
 \item give structure of the report.
 \end{itemize}

\section{The sRNA regulatory system}
\label{The sRNA regulatory system}

 \begin{itemize}
 \item  
 \item

 \end{itemize}

\begin{figure}[H]
\centering
\includegraphics[trim = 60 630 300 30,page=2,clip = true]{pnas1203831109.pdf}
\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
\label{community}
\end{figure}

>>>>

\author{J.~Binysh
        \\ \IEEEmembership{* University of Warwick, Complexity Department}}

@@

\begin{abstract}
genetic circuitry, RNA's offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription. In comparison to proteins, it is relatively straightfor
\end{abstract}

@@

\section{Introduction}
\label{sec: Intro}
The process of gene expression can be briefly summarised as: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}). This RNA molecule (known as messenger RNA, or mRNA) makes its way to a piece of cellular machinery called the Ribosome, which reads it, and makes a protein - which protein is made depends on the DNA sequence originally read (\textit{translation}).

 The path from genetic transcription to protein expression is naturally regulated in many ways \cite{MolecularBiology}. This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues. The natural cell machinery which performs this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.

@@

\section{The sRNA regulatory system}
\label{The sRNA regulatory system}
In bacteria, one mechanism by which gene expression is regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA. This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA. Translation may be self repressed by the tail of the mRNA folding over and binding across the RBS, forming a stem loop in the mRNA and preventing the Ribosome from binding (Fig. \ref{RBS}). This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind (Fig. \ref{RBS}}). In summary, the presence of the sRNA positively regulates gene expresssion.
\todo[inline]{define sRNA}

\begin{figure}[H]
\centering
\includegraphics[trim = 100 170 100 400,page=10,clip = true,scale = 0.6]{Appendix.pdf}
\caption{A mechanism by which sRNA's can regulate gene expression. Initially, the mRNA is folded over on itself in a loop, covering the RBS and blocking Ribosome binding. The sRNA binds to this mRNA, causing a conformational change which uncovers the RBS, and allows translation to occur. }
\label{RBS}
\end{figure}

\cite{Rodrigo2012} proposed a computational methodology to design general genetic circuits based on RNA interactions, and as a case study of the methodology chose to design a synthetic sRNA- mRNA pair capable of acting in the manner described above. The algorithm assumed an interaction scheme between the RNA's as shown in Fig. \ref{reactionscheme}. The two RNA's, originally in their own individually folded states, would initially interact via a small 'toehold' sequence of unpaired nucleotides to form an unstable transition state. This intermediate complex would then rapidly form a final, stable complex with the desired conformation. By suggesting sRNA and mRNA sequences which optimised this energy landscape, \cite{Rodrigo2012} suggested several devices which would form a stable hybrid with the RBS free, and experimentally validated their function in \textit{E. coli}.

\begin{figure}[H]
\centering
\includegraphics[trim = 60 630 300 30,page=2,clip = true]{pnas1203831109.pdf}
\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
\label{reactionscheme}
\end{figure}

Further, by placing the concentrations of the sRNA and mRNA under the control of tuneable promoters, \cite{Rodrigo2012} constructs a logical AND gate from one of the proposed devices (RAJ11) \textit{in vivo} (Fig. \ref{ANDGate}).

\begin{figure*}[b!]
\centering
\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
\label{ANDGate}
\end{figure}
